---
title: "CS112 Assignment 2, Fall 2020"
author: "Jomo Tafari Dixon"
date: "09/27/2020"
output:
  pdf_document: pdf_document
  word_document: default
---

```{r setup, include=FALSE}
#Don't change this part of the document
## install and load the necessary packages
library(lubridate)
library(tree)
library(Matching)
library(boot)
library(randomForest)
# we need to set the seed of R's random number generator, in order to produce comparable results 
set.seed(32)
```

**Note**: *This is an RMarkdown document. Did you know you can open this document in RStudio, edit it by adding your answers and code, and then knit it to a pdf? Then you can submit both the .rmd file (the edited file) and the pdf file as a zip file on Forum. This method is actually preferred. To learn more about RMarkdown, watch the videos from session 1 and session 2 of the CS112B optional class. [This](https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf) is also a cheat sheet for using Rmarkdown. If you have questions about RMarkdown, please post them on Perusall.*

**Note**: *If you are not comfortable with RMarkdown, you can use any text editor (google doc or word) and type your answers and then save everything as a pdf and submit both the pdf file AND the link to your code (on github or jupyter) on Forum.*

**Note**: *Try knitting this document in your RStudio. You should be able to get a pdf file. At any step, you can try knitting the document and recreate a pdf. If you get error, you might have incomplete code.*

**Note**: *If you are submitting your assignment as an RMarkdown file, make sure you add your name to the top of this document.*

## QUESTION 1

#### STEP 1

Create a set of 1000 outcome observations using a data-generating process (DGP) that incorporates a systematic component and a stochastic component (of your choice)

```{r}
# YOUR CODE HERE
set.seed(232234)
q1.var1.t <- sort(rnorm(250, mean = 87, sd = 4), decreasing = FALSE)
q1.var1.c <- sort(rnorm(250, mean = 64, sd = 7), decreasing = FALSE)
q1.var2.t <- sort(rnorm(250, mean = 82, sd = 6), decreasing = FALSE)
q1.var2.c <- sort(rnorm(250, mean = 62, sd = 7.5), decreasing = FALSE)
q1.var3.t <- sort(rnorm(250, mean = 83, sd = 5.5), decreasing = FALSE)
q1.var3.c <- sort(rnorm(250, mean = 61, sd = 8), decreasing = FALSE)
q1.var1 <- sort(cbind(q1.var1.t, q1.var1.c), decreasing = FALSE)
q1.var2 <- sort(cbind(q1.var2.t, q1.var2.c), decreasing = FALSE)
q1.var3 <- sort(cbind(q1.var3.t, q1.var3.c), decreasing = FALSE)
q1.out.var1 <- sort(sample(c(0,1), 500, replace = TRUE), decreasing = FALSE)
q1.dataframe1 <- data.frame(q1.var1, q1.var2, q1.var3, q1.out.var1)
q1.glm1 <- glm(q1.out.var1 ~., q1.dataframe1, family =  binomial)
summary(q1.glm1)

```

#### STEP 2

Tell a 2-3 sentence story about the data generating process you coded up above. What is it about and what each component mean?

The above is a function that takes in test scores of students on 4 seperate exams, and predicts their future score on an IQ test given 



#### STEP 3

Using an incorrect model of the systematic component (of your choice), and using the simulation-based approach covered in class (the `arm` library, etc.), generate 1 prediction for every observation and calculate and report RMSE. Make sure you write out your model and report your RMSE. 

Each prediction should be conditional on the predictor values of the corresponding observation. E.g., when predicting for observation #1, use the predictor values of observation #1.

```{r}
# YOUR CODE HERE

```


#### STEP 4

Using the correct model (correct systematic and stochastic components), and using the simulation-based approach covered in class (the `arm` library, etc.), generate 1 prediction for every observation and calculate & report your RMSE. Once again, write out your model and report your RMSE. 

Each prediction should be conditional on the predictor values of the corresponding observation. E.g., when predicting for observation #1, use the predictor values of observation #1.

```{r}
# YOUR CODE HERE

```



#### STEP 5

Which RMSE is larger: The one from the correct model or the one from the incorrect model? Why?




## QUESTION 2

Imagine that you want to create a data viz that illustrates the sensitivity of regression to outlier data points. So, you want to create two figures: 
	
One figure that shows a regression line fit to a 2-dimensional (x and y) scatterplot, such that the regression line clearly has a positive slope. 

```{r}
# YOUR CODE HERE
set.seed(457345)
q2.var1 <- sort(rnorm(10, mean = 87, sd = 8), decreasing = FALSE)
q2.var2 <- sort(rnorm(10, mean = 63, sd = 12), decreasing = FALSE)
q2.dataframe1 <- data.frame(q2.var1, q2.var2)
q2.lm1 <- lm(q2.var2 ~ q2.var1, q2.dataframe1)
q2.lm1
plot(q2.dataframe1)
abline(q2.lm1)
```

And, another figure that shows a regression line with a negative slope fit to a scatter plot of the same data **plus one additional outlier data point**. This one data point is what changes the sign of the regression line’s slope from positive to negative.

```{r}
# YOUR CODE HERE
set.seed(457345)

q2.var1 <- sort(rnorm(10, mean = 87, sd = 8), decreasing = FALSE)
q2.var2 <- sort(rnorm(10, mean = 63, sd = 12), decreasing = FALSE)
q2.dataframe1 <- data.frame(q2.var1, q2.var2)

q2.func <- function(original, q2.var1.ad, q2.var2.ad){
  new1 <- data.frame(q2.var1.ad, q2.var2.ad)
  new2 <- rbind(original, new1)
  return(new2)
}
q2.dataframe2 <- q2.func(q2.dataframe1, 99, 12)
q2.lm2 <- lm(q2.var2 ~ q2.var1, q2.dataframe2)
plot(q2.dataframe2)
abline(q2.lm1)
```

Be sure to label the axes and the title the figures appropriately. Include a brief paragraph that explains the number of observations and how you created the data set and the outlier.




## QUESTION 3

#### STEP 1

Using the `laLonde` data set, run a linear regression that models `re78` as a function of `age`, `education`, `re74`, `re75`, `hisp`, and `black`. Note that the `lalonde` data set comes with the package `Matching`.

```{r}
# YOUR CODE HERE
library(Matching)
data("lalonde")
lm.q3 <- lm(re78 ~ ., lalonde)
```

#### STEP 2

Report coefficients and R-squared. 

```{r}
coef.rsq <- c(coef(lm.q3), summary(lm.q3)$r.squared)
coef.rsq
```

Then calculate R-squared by hand and confirm / report that you get the same or nearly the same answer as the summary (`lm`) command. 

Write out hand calculations here.
average coef = 5.357093e+01 + 4.007703e+02 + -2.037333e+03 + 4.258185e+02 + -1.463292e+02 + -1.517944e+01 + 1.233988e-01 + 1.974299e-02 + 1.380285e+03 + -1.071215e+03 + 1.670709e+03 = 6.011274834e+01

5.357093e+01 - 6.011274834e+01 = -6.54181834
4.007703e+02 - 6.011274834e+01 = 3.4065755166e+02
-2.037333e+03 - 6.011274834e+01 = -2.09744574834e+03
4.258185e+02 - 6.011274834e+01 = 3.6570575166e+02
-1.463292e+02 - 6.011274834e+01 = -2.0644194834e+02
-1.517944e+01 - 6.011274834e+01 = -7.529218834e+01
1.233988e-01 - 6.011274834e+01
1.974299e-02 - 6.011274834e+01
1.380285e+03 - 6.011274834e+01
-1.071215e+03 - 6.011274834e+01
1.670709e+03 - 6.011274834e+01





#### STEP 3

Then, setting all the predictors at their means EXCEPT `education`, create a data visualization that shows the 95% confidence interval of the expected values of `re78` as `education` varies from 3 to 16. Be sure to include axes labels and figure titles.

```{r}
# YOUR CODE HERE

```

#### STEP 4

Then, do the same thing, but this time for the predicted values of `re78`. Be sure to include axes labels and figure titles.

```{r}
# YOUR CODE HERE

```

#### STEP 5

Lastly, write a short paragraph with your reflections on this exercise (specifically, the length of intervals for given expected vs. predicted values) and the results you obtained.




## QUESTION 4

#### STEP 1

Using the `lalonde` data set, run a logistic regression, modeling treatment status as a function of `age`, `education`, `hisp`, `re74` and `re75`. Report and interpret the regression coefficient and 95% confidence intervals for `age` and `education`.

```{r}
# YOUR CODE HERE

```

Report and interpret regression coefficient and 95% confidence intervals for `age` and `education` here. 



#### STEP 2

Use a simple bootstrap to estimate (and report) bootstrapped confidence intervals for `age` and `education` given the logistic regression above. Code the bootstrap algorithm yourself.

```{r}
# YOUR CODE HERE

```

Report bootstrapped confidence intervals for `age` and `education` here. 




#### STEP 3

Then, using the simulation-based approach and the `arm` library, set all the predictors at their means EXCEPT `education`, create a data visualization that shows the 95% confidence interval of the expected values of the probability of receiving treatment as education varies from 3 to 16. Be sure to include axes labels and figure titles.

```{r}
# YOUR CODE HERE

```

#### STEP 4

Then, do the same thing, but this time for the predicted values of the probability of receiving treatment as education varies from 3 to 16. Be sure to include axes labels and figure titles.

```{r}
# YOUR CODE HERE

```

#### STEP 5

Lastly, write a short paragraph with your reflections on this exercise and the results you obtained.



## QUESTION 5


Write the executive summary for a decision brief about the impact of a stress therapy program, targeted at individuals age 18-42, intended to reduce average monthly stress. The program was tested via RCT, and the results are summarized by the figure that you get if you run this code chunk:

```{r}
# Note that if you knit this document, this part of the code won't 
# show up in the final pdf which is OK. We don't need to see the code
# we wrote.

# How effective is a therapy method against stress

# Participants in the study record their stress level for a month.
# Every day, participants assign a value from 1 to 10 for their stress level. 
# At the end of the month, we average the results for each participant.

#adds the confidence interval (first row of the matrix is lower 
# bound, second row is the upper bound)
trt1 = matrix(NA,nrow=2,ncol=7)
ctrl = matrix(NA,nrow=2,ncol=7) 

trt1[,1]=c(3.7, 6.5) #18  
ctrl[,1]=c(5, 8)

trt1[,2]=c(5, 8.5) #22
ctrl[,2]=c(7.5, 9)

trt1[,3]=c(6, 9) #26
ctrl[,3]=c(8.5, 10)

trt1[,4]=c(5, 7) #30
ctrl[,4]=c(6, 8)

trt1[,5]=c(3.5, 5) #34
ctrl[,5]=c(4.5, 7)

trt1[,6]=c(2, 3.5) #38
ctrl[,6]=c(3.5, 6)

trt1[,7]=c(0.5, 2) #42
ctrl[,7]=c(2.5, 5)

# colors to each group
c1 = rgb(red = 0.3, green = 0, blue = 1, alpha = 0.7) #trt1
c2 = rgb(red = 1, green = 0.6, blue = 0, alpha = 1) #trt2
c3 = rgb(red = 0, green = 0.5, blue = 0, alpha = 0.7) #ctrl

# creates the background of the graph
plot(x = c(1:100), y = c(1:100), 
     type = "n", 
     xlim = c(17,43), 
     ylim = c(0,11), 
     cex.lab=1,
     main = "Stress Level - 95% Prediction Intervals", 
     xlab = "Age", 
     ylab = "Average Stress Level per Month", 
     xaxt = "n")

axis(1, at=seq(18,42,by=4), seq(18, 42, by=4))

grid(nx = NA, ny = NULL, col = "lightgray", lty = "dotted",
     lwd=par("lwd"), equilogs = TRUE)

# adds the legend
legend('topright',legend=c('Treatment','Control'),fill=c(c1,c2))

# iterates to add stuff to plot
for (age in seq(from=18,to=42,by=4)) { 
  #treatment
  segments(x0=age-0.2, y0=trt1[1, (age-18)/4+1],
           x1=age-0.2, y1=trt1[2, (age-18)/4+1], lwd=4, col=c1)
  
  #control
  segments(x0=age+0.2, y0=ctrl[1, (age-18)/4+1],
           x1=age+0.2, y1=ctrl[2, (age-18)/4+1], lwd=4, col=c2)
}

```

(Not that it matters, really, but you can imagine that these results were obtained via simulation, just like the results you have hopefully obtained for question 2 above). 

Your executive summary should be between about 4 and 10 sentences long, it should briefly describe the the purpose of the study, the methodology, and the policy implications/prescription. (Feel free to imaginatively but realistically embellish/fill-in-the-blanks with respect to any of the above, since I am not giving you backstory here).

Write your executive summary here.

This study aims to predict the expected effect of an intervention on those who opt for it’s treatment. Based on our concluded study, those between the ages of 18 and 26 can expect with greater certainty that our drug will indeed have the expected effect on lowering your stress levels.


## QUESTION 6

Can we predict what projects end up being successful on Kickstarter? 

We have data from the [Kickstarter](https://www.kickstarter.com/) company. 

From Wikipedia: Kickstarter is an American public-benefit corporation based in Brooklyn, New York, that maintains a global crowdfunding platform focused on creativity and merchandising. The company's stated mission is to "help bring creative projects to life". As of May 2019, Kickstarter has received more than $4 billion in pledges from 16.3 million backers to fund 445,000 projects, such as films, music, stage shows, comics, journalism, video games, technology, publishing, and food-related projects.

The data is collected by [Mickaël Mouillé](https://www.kaggle.com/kemical) and is last uodated in 2018. Columns are self explanatory. Note that `usd_pledged` is the column `pledged` in US dollars (conversion done by kickstarter) and `usd_pledge_real` is the `pledged` column in real US dollars of the pledged column. Finally, `usd_goal_real` is the column `goal` in real US dollars. You should use the real columns.


So what makes a project successful? Undoubtedly, there are many factors, but perhaps we could set up a prediction problem here, similar to the one from the bonus part of the last assignment where we used GDP to predict personnel contributions. 

We have columns representing the the number of backers, project length, the main category, and the real project goal in USD for each project. 

Let's explore the relationship between those predictors and the dependent variable of interest — the success of a project. 

Instead of running a simple linear regression and calling it a day, let's use cross-validation to make our prediction a little more sophisticated. 

Our general plan is the following: 

1. Build the model on a training data set 
2. Apply the model on a new test data set to make predictions based on the inferred model parameters. 
3. Compute and track the prediction errors to check performance using the mean squared difference between the observed and the predicted outcome values in the test set. 

Let's get to it, step, by step. Make sure you have loaded the necessary packages for this project. 

#### STEP 1: Import & Clean the Data

Import the dataset from this link: https://tinyurl.com/KaggleDataCS112 

Remove any rows that include missing values. 

```{r}
# YOUR CODE HERE

```

#### STEP 2: Codify outcome variable

Create a new variable that is either successful or NOT successful and call it `success` and save it in your dataframe. It should take values of 1 (successful) or 0 (unsuccessful).

```{r}
# YOUR CODE HERE

```

#### STEP 3: Getting the project length variable  

Projects on Kickstarter can last anywhere from 1 - 60 days. Kickstarter claims that projects lasting any longer are rarely successful and campaigns with shorter durations have higher success rates, and create a helpful sense of urgency around your project. Using the package `lubridate` or any other package in R you come across by Googling, create a new column that shows the length of the project by taking the difference between the variable `deadline` and the variable `launched`. Call the new column `length` and save it in your dataframe.

Remove any project length that is higher than 60. 

```{r}
# YOUR CODE HERE

```

#### STEP 4: Splitting the data into a training and a testing set

While there are several variations of the k-fold cross-validation method, let’s stick with the simplest one where we just split randomly the dataset into two (k = 2) and split our available data from the link above into a training and a testing (aka validation) set. 

Randomly select 80% of the data to be put in a training set and leave the rest for a test set. 

```{r}
# YOUR CODE HERE

```


#### STEP 5: Fitting a model 

Use a logistic regression to find what factors determine the chances a project is successful. Use the variable indicating whether a project is successful or not as the dependent variables (Y) and number of backers, project length, main category of the project, and the real project goal as independent variables. Make sure to use the main category as factor.

```{r}
# YOUR CODE HERE

```


#### STEP 6: Predictions

Use the model you’ve inferred from the previous step to predict the success outcomes in the test set.

```{r}
# YOUR CODE HERE

```

#### STEP 7: How well did it do? 

Report the Root Mean Squared Error (RMSE) of the predictions for the training and the test sets. 

```{r}
# YOUR CODE HERE

```

#### Step 8: LOOCV method

Apply the leave-one-out cross validation (LOOCV) method to the training set. What is the RMSE of the training and test sets. How similar are the RMSEs?

```{r}
# YOUR CODE HERE

```


#### Step 9: Explanations

Compare the RMSE from the simple method to the LOOCV method?

How do data scientists really use cross-validation? How is the approach in this project differ from real-world cases? Give an example to make your point!




## Extra Credit: Least Absolute Deviation Estimator

#### STEP 1

Figure out how to use rgenoud to run a regression that maximizes the least absolute deviation instead of the traditional **sum of the squared residuals**. Show that this works by running that regression on the `lalonde` data set with outcome being `re78` and independent variables being `age`, `education`, `hisp`, `re74`, `re75`, and `treat`. 

```{r}
# YOUR CODE HERE

```


#### STEP 2

How different is this coef on treat from the coef on treat that you get from the corresponding traditional least squares regression?





#### STEP 3

Now figure out how to do the same by using rgenoud to run the logistic regression (modeling treatment status as a function of `age`, `education`, `hisp`, `re74` and `re75`).

```{r}
# YOUR CODE HERE

```


## END OF Assignment!!!

## Final Steps

### Add Markdown Text to .Rmd

Before finalizing your project you'll want be sure there are **comments in your code chunks** and **text outside of your code chunks** to explain what you're doing in each code chunk. These explanations are incredibly helpful for someone who doesn't code or someone unfamiliar to your project.
You have two options for submission:

1. You can complete this .rmd file, knit it to pdf and submit both the .rmd file and the .pdf file on Forum as one .zip file.
2. You can submit your assignment as a separate pdf using your favorite text editor and submit the pdf file along with a lint to your github code. Note that links to Google Docs are not accepted.


### Knitting your R Markdown Document

Last but not least, you'll want to **Knit your .Rmd document into an HTML document**. If you get an error, take a look at what the error says and edit your .Rmd document. Then, try to Knit again! Troubleshooting these error messages will teach you a lot about coding in R. If you get any error that doesn't make sense to you, post it on Perusall.

### A Few Final Checks

If you are submitting an .rmd file, a complete project should have:

- Completed code chunks throughout the .Rmd document (your RMarkdown document should Knit without any error)
- Comments in your code chunks
- Answered all questions throughout this exercise.

If you are NOT submitting an .rmd file, a complete project should have:

- A pdf that includes all the answers and their questions.
- A link to Github (gist or repository) that contais all the code used to answer the questions. Each part of you code should say which question it's referring to.